Vragenlijst thesis evaluatie

### 2015-06-08 doc/template/chap-5.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 1				%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What is cyber security?}

Cyber security is a part of security that focuses on securing computer controlled devices such as computers and smartphones, as well as public and private computer networks, including the entire Internet. A private network can be an enterprise network that is protected from the public network such as the internet.  The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity). These are the 3 key attributes of security, also known as the CIA triad. \\

In order to keep a system secure, it is important to mitigate the possible threats. A threat refers to anything that can breach the security and that can cause a possible harm. The some of the most common threats are spam, malware, spoofing, phishing and  DDoS attacks. 
The biggest threat are caused by malware. 
This is a piece of malicious program that is designed to penetrate unprotected systems or computers, and getting there to sensitive information. Malware can be divided into different subclasses. These subclasses are not mutually exclusive:

\begin{description}
\item Virus: This is a malicious piece of code that replicates and tries to spread to infect other systems or files. The "I love you" virus is an example of a virus that quickly spreads. This virus propagates via mail systems. If someone opens an email with "I love you" with the virus in the annex this virus spreads itself by sending a mail to everyone in your contact list. So the virus can multiply rapidly and eventually a business network might shut down by the heavy traffic. In this example, there is a need for human interaction to spread the virus. If no one opens the mail the virus can not infect other systems.
Unfortunately, there are also viruses that can spread without human interaction. These viruses are referred to as worms. A worm is also a computer program that replicates itself to spread to other computers. Via a computer network, copies of the worm can be forwarded without an intermediary. The worm will use vulnerabilities of the system to infect other computers.
The Stuxnetworm is a very famous worm. Initially this worm spread via infected USB sticks and from then it could spread through the Internet to other computers. The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors. Many reactors have been infected. From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread. 
\item Rootkit:
\item Trojan:
\item Backdoors:
\end{description}

A security report of 2014 reveales that 61\% of the attacks on companies are caused by malware. A company can take different measures to defend himself against malware. 
The threats caused by malware can be devided into 3 big groups: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) [\todo{toevoegen referentie naar kaspersky APT report}].

The known threats easy to defend against. Unkown also. Then the remaining 1\% are the advanced threats. These threats are also known as APT. APT's are ... 
In this paper we focus on targeted attacks. Targeted attacks are malware designed to attack exclusively an organisation. Some examples of the biggest most rare APTs  to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}]

An APT is a persistent targeted attack that tries to penetrate a network, harm and stays unseen for a long period of time. The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other ideological attacks. APT stands for Advanced persistant threat. Advanced stands for the fact that these attacks are well funded and that the attacker itself has a great expertise. These APT are not easy to protect against. Not all APTs are technical advanced. The attacker can also try to exploit existing vulnerabilities and hope that his target has not yet secured himself against these vulnerabilities. Persistent is the fact that the attacker keeps on trying to break in to his victim. He will not give up. The attack can be over various years and different steps. The threats stands for the fact that an APT is ..

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 1				%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This chapter provides the reader with an introduction to the general context of the work presented in this paper.  Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threaths that are in the scope for this work. Section 2.2 then introduces the reader into the main principles of game theory. Subsequently, section 2.3 introduces the reader to the FlipIt game, the specific game that will be used to model cyber security attacks of periodic nature and including a delay.  Finally, section 2.4 gives an overview of the related work, and how the research presented in this thesis is positioned compared to existing results.

\section{What is cyber security?}

Before the digitalization of documents, information was kept on paper and the security of this information was ensured by administrative and physical means. For example, you needed a key to access documents stored in a room full of cabinets where the files were kept. In today's digital era more and more information is kept in a digital format, stored on a computer.  As digitalization progressed, the need for ensuring the security of digital information arose and automated tools where developed for protecting files stored on a computer.  The generic name to protect data stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security. \\

Security is a general term that encompasses several dimensions. More specifically, computer security has tree key objectives that are fundamental to computer security:
\begin{description}
\item Confidentiality: This assures that the confidential of private data is not disclosed or made available to users that do not have the authorization. 
\item Integrity: This assures that data can not be altered by an unauthorized individual.
\item Availability: This assures that data is always accessible and that the service is not denied to authorized individuals.
\end{description}

 the purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).

 These are the 3 key attributes of security, also known as the CIA triad. The tree concepts are fundamental security objectives for the securing of data, information and computing services. \\

''Cyber security is the process of applying security tools to ensure confidentiality, integrity, and availability of data. Cybersecurity attempts to ensure the protection of assets, which includes data, desktops, servers, buildings, and most importantly, humans. The goal of cyber security is to protect data both in transit and at rest. Countermeasures can be put in place in order to increase the security of data. Some of these measures include, but are not limited to, access control, awareness training, audit and accountability, risk assessment, penetration testing, vulnerability management, and security assessment and authorization.'' \todo{tekst overgenomen van wikipedia} \\

In the context of cyber security, the terms ' threat and attack' are often used interchangeably, referring to more or less the same thing. They however have a slightly different meaning. A threat refers to anything that can breach the security and that can cause a possible harm. It is a possible danger that can exploit a vulnerability. An attack is an assault on computer security that comes from a threat. It is an intelligent act that tries deliberately to breach security through vulnerabilities.  \\

In order to keep a system secure, it is important to mitigate the possible threats. The possible threats can take many forms, the most commen being spam, malware, spoofing, phishing and  DDoS attacks. 
A security report of 2014 reveals that 61\% of the attacks on companies are caused by malware. This is a piece of malicious program that is designed to penetrate unprotected systems or computers, and getting there to sensitive information. Malware can be divided into different subclasses that are not mutually exclusive:

\begin{description}
\item Virus: This is a malicious piece of code that replicates and tries to spread to infect other systems or files. The "I love you" virus is an example of a virus that quickly spreads. This virus propagates via mail systems. If someone opens an email with "I love you" with the virus in the annex this virus spreads itself by sending a mail to everyone in your contact list. So the virus can multiply rapidly and eventually a business network might shut down by the heavy traffic. In this example, there is a need for human interaction to spread the virus. If no one opens the mail the virus can not infect other systems.
Unfortunately, there are also viruses that can spread without human interaction. These viruses are referred to as worms. A worm is also a computer program that replicates itself to spread to other computers. Via a computer network, copies of the worm can be forwarded without an intermediary. The worm will use vulnerabilities of the system to infect other computers.
The Stuxnetworm is a very famous worm. Initially this worm spread via infected USB sticks and from then it could spread through the Internet to other computers. The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors. Many reactors have been infected. From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread. 
\item Rootkit: A rootkit is designed to be stealthy and hide a set of programs installed on a system from methods of detection.  The programs have administrator or root privileges to that system, which makes it possible to access the functions and services of the operating system, change files, take control over the monitor processes and send and receive network traffic.  A rootkit can make changes to the system to keep itself concealed from detection.
\item Trojan: This is a malicious program that disguises itself as something normal, so that users won't be suspicious of installing it.  
\item Backdoors: Also know as trapdoor, is a whole in the system or program that allows access 
\end{description}

A company can take different measures to defend itself against malware. The threats caused by malware can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy [\todo{toevoegen referentie naar kaspersky APT report}].

The known threats are the easiest to  defend oneself against. Standard malware protection tools like firewalls and virus scanners can keep these kind of malware out of the system. Installing protection against unkown is also relatively easy. Then the remaining 1\% are the advanced threats, also known as APT. APT's are ... 
In this paper we focus on targeted attacks. Targeted attacks are malware designed to attack exclusively an organisation.

An APT is a persistent targeted attack that tries to penetrate a network to harm it while staying unseen for a long period of time. The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or some other kind of ideological attacks. Advanced Persistent Threat are called  'Advanced' for the fact that these attacks are well funded and that (usually) the attacker itself needs a great expertise to successfully penetrate a network. Not all APTs are technical advanced though. The attacker can also try to exploit existing vulnerabilities simply based in the hope that his target has not yet secured himself against these vulnerabilities. 'Persistent' refers to the fact that the attacker keeps on trying to break in to his victim. He will not give up. The attack can be over various years and different steps. The threats stands for the fact that an APT is attempting to break the protection of valuable data the leakage of which outside the secure environment may severely damage the organisation owning the data.

Some examples of the biggest most rare APTs are listed to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}]
